---
title: "BAC Logistic Regression"
output: html_notebook
---
 
##Load Libraries
```{r}
library(modelr)
library(tidyverse)
library(descriptr)
install.packages("margins")
library(margins)
install.packages("gmodels")
library(gmodels)
```

##Import Data
```{r}
obs <- read_csv("bac_obs.csv")
```


```{r}
male <- 7/10
no_male <- 3/10

female <- 3/10
no_female <- 7/10

odd_male <- 7/3
odd_female <- 3/7

OR_male <- odd_male/odd_female
OR_female <- odd_female/odd_male

```






##Dichotomize two of the variables for demonstration purposes
```{r}
# Create dichotomized versions of bac (>.08 vs. <= .08) and
# typ_drks (> average of 2 per day vs. <= average of 2 per day).
obs <- mutate(obs,
bac_over = ifelse(bac > .08, 1, 0),
typ_hvy = ifelse(typ_drks > 60, 1, 0))
```

##2x2 Crosstab
```{r}
#ds_cross_table(obs$typ_hvy, obs$bac_over)
CrossTable(obs$typ_hvy, obs$bac_over)
```

##Calculate proportions for BAC categories by typical drinking categories
```{r}
# calculate proportion of light drinkers and heavy drinkers who had a BAC at or under .08
prop_light_under <- 90/161
prop_heavy_under <- 3/39
# calculate proportion of light drinkers and heavy drinkers who had a BAC over .08
prop_light_over <- 71/161
prop_heavy_over <- 36/39
```

##Risk Ratio
```{r}
RR = prop_heavy_over/prop_light_over
RR
```

##Probability to Odds
```{r}
# calculate the odds of BAC being over .08 for heavy drinkers
odds_heavy <- prop_heavy_over/prop_heavy_under
odds_heavy
# calculate the odds of BAC being over .08 for light drinkers
odds_light <- prop_light_over/prop_light_under
odds_light
```

##Transofmration of Probability to Odds
```{r}
trans_probodds <- tibble(prob = seq(0, 1, by = .05)) %>%
mutate(odds = prob/(1-prob))
```

##Calcuate the Odds Ratio to Compare the Odds Across Two Groups
```{r}
# calculate the odds ratio
OR <- odds_heavy/odds_light
OR
# reverse the numerator and denominator
OR_rev <- odds_light/odds_heavy
OR_rev
```


##Calculate the proportion of people at each level of alc_gm who had a BAC over .08
```{r}
avg_over_bygrams <- obs %>%
group_by(alc_gm) %>%
summarize(proportion_over = mean(bac_over)) %>%
ungroup()
```

##Plot the mean BAC across levels of alc_gm
```{r}
ggplot(avg_over_bygrams, aes(y = proportion_over, x = alc_gm)) +
geom_point() +
labs(title = "Proportion of women with BAC over .08 by alcohol consumed",
x = "Grams of alcohol consumed", y = "Proportion of women with BAC over.08")
```

#The Logistic Regression Model
##Logistic Regression for a Continuous Predictor
```{r}
# first center alc_gm at the mean
obs <- mutate(obs, alc_gm_m = alc_gm - mean(alc_gm))
# fit the logistic regression model
logreg0 <- glm(bac_over ~ alc_gm_m, data = obs, family=binomial("logit"))
summary(logreg0)

##INTERPRET THIS IN THE LOG OG THE ODDS! PREDICTED Y = LOG OF ODDS!
```


##Significance Test
```{r}
confint(logreg0, level = .95)
```


##Transform Log Odds Coefficients Back to Odds
###Odds Ratio
```{r}
exp(coef(logreg0))
```


###Transform log odds coef to percent change
```{r}
# function to interpret regression slope for ln transformed y, original x
ln.y <- function(slope, x_chg) {
new_slope <- 100 * (exp(slope * x_chg)-1)
return(new_slope)
}
ln.y(slope = .44883, x_chg = 1)
```



###95% Confidence Intervals for Exponentiated Coefficients
```{r}
exp(cbind(OR = coef(logreg0), confint(logreg0)))
```

##Obtain Predicted Values Based on Model Estimates
```{r}
# predicted value when alc_gm is at the mean (alc_gm_m = 0)
logodds <- 0.44099 + (0.44883*0)
odds <- exp(logodds)
prob <- odds/(1+odds)
logodds
odds
prob
```

##Get predicted values using code
```{r}
pred_grid0 <- tibble(alc_gm_m = 0) %>%
add_predictions(logreg0) %>%
mutate(odds = exp(pred),
predprob = (odds/(1+odds)))
pred_grid0
```

##Plot of the Fitted Model
```{r}
pred_grid0 <- data_grid(obs, alc_gm_m = seq_range(alc_gm_m, 10)) %>%
add_predictions(logreg0) %>%
mutate(odds = exp(pred),
predprob = (odds/(1+odds)),
alc_gm = alc_gm_m + mean(obs$alc_gm))
```

##Plot in the metric of log odds
```{r}
ggplot(pred_grid0, aes(x = alc_gm, y = pred)) +
geom_line(size = 1) +
labs(title = "Model fitted log odds of BAC > .08 as function of grams of alcohol consumed",
x = "Grams of Alcohol Consumed", y = "Log odds of BAC > .08")
```


##Plot in the mdetric of odds
```{r}
ggplot(pred_grid0, aes(x = alc_gm, y = odds)) +
geom_line(size = 1) +
labs(title = "Model fitted odds of BAC > .08 as function of grams of alcohol consumed",
x = "Grams of Alcohol Consumed", y = "Odds of BAC > .08")
```


##Plot in Metric of Probability
```{r}
ggplot(pred_grid0, aes(x = alc_gm, y = predprob)) +
geom_line(size = 1) +
labs(title = "Model fitted probabilities of BAC > .08 as function of grams of alcohol consumed",
x = "Grams of Alcohol Consumed", y = "Probability of BAC > .08")
```

##Fit a logistic regression model to 2x2 example
```{r}
logreg1 <- glm(bac_over ~ typ_hvy, data = obs, family=binomial("logit") )
summary(logreg1)
exp(cbind(OR = coef(logreg1), confint(logreg1)))
```

##Obtain log odds, odds, and probability
```{r}
pred_grid1 <- data_grid(obs, typ_hvy) %>%
add_predictions(logreg1) %>%
mutate(odds = exp(pred),
predprob = (odds/(1+odds)))
```


##Add additional predictor to our 2x2 example
```{r}
# center alcexp at the mean
obs <- mutate(obs, alcexp_m = alcexp - mean(alcexp))
#fit the regression model
logreg2 <- glm(bac_over ~ typ_hvy + alcexp_m, data = obs, family=binomial("logit"))
summary(logreg2)
exp(cbind(OR = coef(logreg2), confint(logreg2)))
```


##Predicted Scores and Plots
```{r}
pred_grid2 <- data_grid(obs, typ_hvy, alcexp_m = seq_range(alcexp_m, 10)) %>%
add_predictions(logreg2) %>%
mutate(odds = exp(pred),
predprob = (odds/(1+odds)),
typ_hvy = factor(typ_hvy, levels = c(0,1), labels = c("light drinker", "heavy drinker")),
alcexp = alcexp_m + mean(obs$alcexp))
```

##Plot the Results
```{r}
ggplot(pred_grid2, aes(x = alcexp, y = pred, group = factor(typ_hvy), colour = typ_hvy)) +
geom_line(size = 1) +
guides(color=guide_legend("Type of Typical Drinking")) +
labs(title = "Model fitted log odds as function of alcohol expectancies and typical drinking",
x = "Alcohol Expectancies", y = "Log odds of BAC > .08")
```

##Make a Plot of th emodel results - odds
```{r}
ggplot(pred_grid2, aes(x = alcexp, y = odds, group = factor(typ_hvy), colour = typ_hvy)) +
geom_line(size = 1) +
guides(color=guide_legend("Type of Typical Drinking")) +
labs(title = "Model fitted odds as function of alcohol expectancies and typical drinking",
x = "Alcohol Expectancies", y = "Odds of BAC > .08")
```

##Make a plot of the results - the probability
```{r}
ggplot(pred_grid2, aes(x = alcexp, y = predprob, group = factor(typ_hvy), colour = typ_hvy)) +
geom_line(size = 1) +
guides(color=guide_legend("Type of Typical Drinking")) +
labs(title = "Model fitted probabilities as function of alcohol expectancies and typical drinking",
x = "Alcohol Expectancies", y = "Predicted probability of BAC > .08")
```

##Obtain the Marginal effect, choose values of x
```{r}
# make typ_hvy a factor
obs <- mutate(obs, typ_hvy.f = factor(typ_hvy, levels = c(0,1), labels = c("light", "heavy")))
# fit model to feed to margins (don’t center any variables)
demo_margins <- glm(bac_over ~ typ_hvy.f + alcexp, data=obs, family=binomial("logit"))
# marginal effects at chosen levels of the covariates
pick <- margins(demo_margins, at = list(typ_hvy.f = c("light", "heavy"), alcexp = c(2:6)))
summary(pick)
```

##Average marginal effects
```{r}
ame <- margins(demo_margins)
summary(ame)
```

#Interactions in Logistic 
##Fit logistic regression model with interaction
```{r}
# center the predictors
obs <- mutate(obs,
typ_drks_m = typ_drks - mean(typ_drks),
pmood_m = pmood - mean(pmood),
weight_m = weight - mean(weight),
absorb_m = absorb - mean(absorb))
# fit the model
logint <- glm(bac_over ~ typ_drks_m + pmood_m + typ_drks_m*pmood_m + weight_m + absorb_m, data = obs,
family = binomial("logit"))
summary(logint)
# get OR and CIs
exp(cbind(OR = coef(logint), confint(logint)))
```

##Plot the Model Results
```{r}
sd(obs$pmood)
pred_grid_int <- data_grid(obs,
typ_drks_m = seq_range(typ_drks_m, 10),
pmood_m = c(-1.396973, 0, 1.396973),
weight_m = 0,
absorb_m = 0) %>%
add_predictions(logint) %>%
mutate(odds = exp(pred),
predprob = (odds/(1+odds)),
typ_drks = typ_drks_m + mean(obs$typ_drks),
pmood.f = factor(pmood_m, levels = c(-1.396973, 0, 1.396973),
labels = c("low partying mood", "average partying mood", "high partying mood")))
```

##Plot the model in terms of log odds
```{r}
ggplot(pred_grid_int, aes(x = typ_drks, y = pred, group = pmood.f, colour = pmood.f)) +
geom_line(size = 1) +
guides(color=guide_legend("Degree of Partying Mood")) +
labs(title = "Model fitted log odds as function of typical drinking and partying mood",
subtitle = "Weight and absorption are held constant at the mean",
x = "Typical Drinking", y = "Log odds of BAC > .08")
```





#Intro to Additive Interactions in Logistic Regression