---
title: "Psy 350: HW 1"
output: 
  html_notebook:
    toc: true
    toc_float: true
---
#Cleare environments
```{r}
rm(list=ls())
```

#load libraries
```{r}
library(psych)
```

#Read in data
```{r, warning = FALSE}
setwd("C:/Users/Neil/Desktop/Class folder/Fall 2018/Psy_350/Data")

Yetz.data <- read.csv("PSY 350 Class Survey.csv", header = TRUE, stringsAsFactors = FALSE)

```




#1

##a.	

What proportion of the class identifies as male, and what proportion identifies as female? What proportion did not identify as either male or female?  Note that I am asking for proportions here – decimal values – not percentages or numbers of students.

```{r}

prop.table(table(Yetz.data$gender))

```

Male = 16%
Female = 81%

#2

```{r}
table(Yetz.data$interest)
```

##a

How many students in the class are primarily interested in counseling psychology?  Here, I am asking you for a number

Counseling = 27

##b.	
What are the three most popular areas of interest in psychology among the class? 

Top = Counseling, forensic, industrial


##c.

What percentage of the class is still undecided about their primary interest area (that is, selected “unsure”)? 

unsure = 
```{r}
prop.table(table(Yetz.data$interest))
```

unsure = 7%

##d.	

What R function did you use to calculate the previous answer?

prop.table()


#3

##a.	

Create a figure that represents the interest areas of the class. Take a screen shot and upload it.

```{r, fig.height= 12, fig.width = 12}

bar <- table(Yetz.data$interest)

barplot(bar)

```


##b.	

Is it more appropriate here to use a bar chart or a histogram? 

barchart

##c.	

Why?

Nominal Variables


#4

##a.	

What is the mean number of cups of coffee consumed by members of the class? 


```{r}
describe(Yetz.data$coffee)
```

Mean = 4.15


##b.

What is the median? 

Median = 3

##c.

What is the standard deviation for this variable? 

SD = 4.84


#5

##a.	

Create an appropriate figure to represent the number of cups of coffee consumed by members of the class. Take a screen shot and upload it. 

```{r}
hist(Yetz.data$coffee)
```



##b.	

Is it more appropriate here to use a bar chart or a histogram? 

Histogram


#6.

##a.	

Is the distribution of coffee consumption symmetric or skewed, based on the statistics and decision rules discussed in lecture and your text? 

Symmetric: 

##b.	

Explain how you know (report the numbers that support your conclusion). (1 point)


skew = 1.8 less than agreed 2+ for too much skew
