---
title: "R Notebook: Lab 10 Neil Yetz & Ashlie Johnson"
output: 
  html_notebook:
    toc: yes
    toc_float: true
---

#Load Libraries
```{r, message = FALSE}
library(tidyverse)
library(lavaan)
#install.packages("semTools")
library(semTools)
```

#Read in data
```{r, message = FALSE}
items <- read_csv("Spring Fever Raw Data.csv")

items <- items %>% 
  select(-X1)
```

#set model
```{r}

model.2f <-'f1 =~ NA*V1 + V2 + V5 + V6 + V8 + V9 + V12
  f2 =~ NA*V3 + V4 + V7 + V10 + V11
  f1 ~~ 1*f1
  f2 ~~ 1*f2'
```

#Run CFA
```{r}
cfa_1 <- cfa(model = model.2f, data = items)
summary(cfa_1, fit.measures=TRUE, standardized = TRUE)
```

#Residual correlations
```{r}
resid(cfa_1, type = "cor")
```

#Reliability
```{r}
reliability(cfa_1)
```

#Model without item 12
##Model
```{r}

model.2f2 <-'f1 =~ NA*V1 + V2 + V5 + V6 + V8 + V9
  f2 =~ NA*V3 + V4 + V7 + V10 + V11
  f1 ~~ 1*f1
  f2 ~~ 1*f2'
```

##Run CFA
```{r}
cfa_2 <- cfa(model = model.2f2, data = items)
summary(cfa_2, fit.measures=TRUE, standardized = TRUE)
```

##Residual correlations
```{r}
resid(cfa_2, type = "cor")
```
##Reliability
```{r}
reliability(cfa_2)
```

#Lab Exercise Response


1.) Looking at the model fit we saw problems with the CFA on this scale. Specifically we say a very chi-square value of 518.934, a high RMSEA of 0.186 90% CI [0.171, 0.200], a CFI of 0.695, & a TLI of 0.627. All of these values inidcate a bad model fit. Next, we looked at the standardized residual correlations between items. From this correlation table we identified that items 3, 4, 7, 10, & 11 had high residual correlations with eachother.  

We looked at items 3, 4, 7, 10, & 11 and and realized that they have a lot in common. These items differed from the other items because they reflect actual effects of "Spring Break Fever" whereas the other items seem to be more indicative of *feelings* associated wit Spring break fever. 

Based on this we decided to conduct a 2 factor CFA model with these items as our 2nd factor. This was shown to be a better model with a chi-square value of 115.295, a CFI of 0.940, a TLI of 0.925, & a RMSEA of 0.069 90% CI [0.051, 0.086]. The omega coefficient reliability was 0.878. Thus indicating a good reliability.

When analyzing the 2nd, improved model, we did still see a high residualized correlation of items 5 & 8. In our discussion, we dicided to keep them because they seem to be tapping into slightly different areas of the construct. Additionally, the overall model fits fit well. 

However, we then noticed that there was a problem with item 12. It had a low factor loading and the item seemed irrelavant to the construct of "Spring Break Fever". Therefore, we re-ran the analysis without item 12. These are the reported numbers: Chi-Square = 105.769, CFI = 0.938, TLI = 0.921, & the RMSEA = 0.076 90% CI = [0.058, 0.095]. Therefore, because the model fit remained good and our new reliabilty was 0.885. All of our statistics appeared unharmed and our factor loadings were better. 






